### 2025-08-05 project-defense/Presentation.tex
small changes tried some euclid codes

\documentclass{beamer}

\usetheme{Berlin}

\usepackage[utf8]{inputenc}
\usepackage{graphicx}  
\usepackage{amsmath}   
\usepackage{hyperref}  
\usepackage{url}       
\usepackage{cite}
\usepackage{booktabs}

\title{Working Principles of Proof Assistants and Formalization of Some Proofs in Agda}
\author{Ashwot Acharya, Bishesh Bohora, Supreme Chaudary \\Supervisior: K.B Manandhar}

\institute{Kathmandu University}
\date{\today}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
\end{frame}

\section{Introduction}

\begin{frame}{Problem Statement}
    As the title suggests, our project will revolve around exploration of theoretical foundation behind Proof Assistants and practice them.

\end{frame}
\begin{frame}{Motivation}
    \begin{itemize}
  \item Strong interest in mathematics and formal reasoning.
  \item Discovered type theory through internet memes on category theory.
  \item Fascinated by the Four Color Theorem and its computer-assisted proof.
  \item The rise of AI raised the question: \textit{"How do computers understand reasoning?"}
  \item Drawn to functional programming, which closely mirrors mathematical logic and structure.

\end{itemize}

\end{frame}
\begin{frame}{Proof Assistant}
        In computer science and mathematical logic, a proof assistant or interactive theorem prover is a software tool to assist with the development of formal proofs by human–machine collaboration.\cite{wikipedia_proof_assistant}

        Examples:
        \begin{itemize}
            \item Coq
            \item LEAN 
            \item Agda
        \end{itemize}
\end{frame}
\begin{frame}{History}
\begin{itemize}
  \item \textbf{Gödel’s Incompleteness Theorems (1930s)}: Revealed limitations of formal systems; sparked interest in formal logic and verification.

  \item \textbf{ Computability Theory (1940s–50s)}: Turing machines and $\lambda$-calculus laid the groundwork for mechanized reasoning.
\item \textbf{Logic Theorist (1954)}: First automated theorem prover by Newell and Simon, capable of proving theorems in propositional logic.

  \item \textbf{LISP (1960)}: A symbolic programming language created by John McCarthy; became essential for early theorem proving systems.

   \item \textbf{Automath (1967)}: First system to check mathematical proofs using dependent types.

\end{itemize}
\end{frame} 
\begin{frame}

\begin{itemize}
   \item \textbf{LCF \& ML (1970s)}: Introduced tactic-based proofs and the ML programming language; foundational to later systems.

  \item \textbf{Coq (1986)}: A proof assistant based on constructive type theory, supporting verified programming and formal proofs.

  \item \textbf{Isabelle (1989)}: Generic theorem prover with support for multiple logics and strong automation tools.

  \item \textbf{Four-Color Theorem (1996)}: First major mathematical theorem re-verified by proof assistants (Coq and HOL).

  \item \textbf{Feit-Thompson Theorem (2012)}: Large-scale group theory proof formalized in Coq, showcasing proof assistant capability.

  \item \textbf{Lean (2015-2023)}: Modern proof assistant combining type theory with performance and usability; popular in formal math via \texttt{mathlib}.

\end{itemize}
\end{frame}

\section{Glance At Theoretical Foundations}
\begin{frame}{Type Theory}
\begin{itemize}
  \item Type theory is a formal system that classifies expressions by their "types."
  \item Originally developed as an alternative to set theory for foundations of mathematics.
  \item Predecessor to Dependent Type Theory, Martin Löf Type Theory which form basis for various proof assistants.
  \item Types prevent logical paradoxes and provide a basis for constructive reasoning.
\end{itemize}
\end{frame}
\begin{frame}{Curry–Howard Correspondence}
\begin{itemize}
  \item A deep analogy between **logic and computation**:
  \begin{itemize}
    \item Propositions $\leftrightarrow$ Types
    \item Proofs $\leftrightarrow$ Programs
  \end{itemize}
  \item A proof of a proposition is a program of a corresponding type.
  \item Enables writing code that is **correct-by-construction**.
  \item Fundamental to systems like Coq, where proving a theorem is like writing a program.
\end{itemize}
\end{frame}
\begin{frame}{$\lambda$-Calculus and Functional Programming}
    \begin{itemize}
        \item \textbf{Lambda Calculus}: A minimal formal system for function definition and application; the foundation of computation theory.
  \item \textbf{Functional Programming}: Directly inspired by lambda calculus; treats computation as evaluation of mathematical functions.

  \item In proof assistant, Core logic is based on typed lambda calculus.
      \item Tools like Coq and Agda embed functional languages with type theory.
    \end{itemize}
\end{frame}

\section{Methodology and Tool}
\begin{frame}{Methodology}
    \begin{itemize}
      \item Investigating the Coq and the agda proof assistant
      \item Assessing the logic behind each of these proof assistants
      \item Investigating of formalization of proofs in Agda
      \item Dissection of verification process
    \end{itemize}
\end{frame}
\begin{frame}{Agda}
Agda is a functional programming language with dependent types. It is based on Martin Löf Type Theory. And most importantly it is a proof assistant.
\cite{inproceedings}
\end{frame}
\begin{frame}{Why Agda?}
  \begin{itemize}
    \item Dependently Typed programming language
    \item Fully embraces the Curry–Howard isomorphism
    \item More comprehensible code
    \item Active Community
  \end{itemize}
\end{frame}

\section{Work Plan}
\begin{frame}{Work Plan}
  \begin{center}
    \begin{tabular}{p{1cm} p{8cm}}
        \hline
        Week & Work plan \\
        \hline
         1 & Understanding (Dependent) Type Theory and Proof Theory\\
        \hline
         2 & Understanding the implementation of type theory models in digital proof assistant \\
        \hline
         3 & Understanding Purely functional Programming paradigm and $ \lambda $-calculus\\
        \hline
         4 & Working Principles of Agda and its core implementation \\
        \hline 
         5 & Formalization of some proofs in Agda \\
        \hline

    \end{tabular}
\end{center}
\end{frame}

\section{Significance and Expected Outcomes}
\begin{frame}{Significance}
  \begin{itemize}
    \item Writing proofs in a formal language introduces more rigor and redcues ambiguity, and makes it executable allowing computer verification.
    \item Simplifies verification process and makes it error free.
     \item Formalizing some proofs can be difficult, this creates a challange which gives rise to novel ideas.
    \item Opportunity to delve into Constructivism.
    \item Can use same theory to check correctness of computer programs/software which helps in writing secure code. 
  \end{itemize}

\end{frame}
\begin{frame}{Expected Outcomes}
  \begin{itemize}
    \item Improved understanding of proofs, logic and programming.
    \item New Formalization of Proofs of various fields in Agda 
    \item Proofs for Various Algorithms
    \item (Re)Discovery of limitations of such proof assistants and formal language.
    \item Future prospects and discussion on possibility of integration of formal language in Machine Learning,
  \end{itemize}
\end{frame}

\section{References}
\begin{frame}{References}
\bibliographystyle{apalike}       
\bibliography{references} 
\end{frame}

\end{document}