### 2025-08-05 final-presentation/final.tex
final presentation kYOUR

\documentclass[aspectratio=169, compress, 12pt]{beamer}

\usepackage[english]{babel}                 %%%%%%%%%%%%%%%%%% (EDIT LANGUAGE)

\usepackage{Custom}                         %%%%%%%%%%%%%%%%%%    (EDIT NAMES)
\usepackage{Custom-2}                         %%%%%%%%%%%%%%%%%%    (EDIT COLOURS)
\usepackage{AuxiliaryFiles/AuxiliaryFiles}  %%%%%%%%%%%%%    (DO NOT EDIT)

\begin{document}
\TitlePage        
\SectionPage    
\SubsectionPage    %%%% PAGE AT THE BEGINNIN OF A SECTION - IF NOT NEEDED, COMMENT IT
\ProgressBar       %%%% PROGRESS BAR ON THE RIGHT MARGIN - IF NOT NEEDED, COMMENT IT
\PageNumbering     %%%% SLIDES NUMBERING - IF NOT NEEDED, COMMENT IT

\input{AuxiliaryFiles/SettingBackup}

\input{AuxiliaryFiles/Thanks}  
\end{document}